---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Create a directory for the created figures
Create directory for the downloaded file if it did not exists
Download zipfile and unzip it
load het dataset
Default language to english
```{r}
Sys.setenv(LANG = "en")
setInternet2(TRUE)

if(!file.exists("figure")){
  dir.create("figure");
}

if(!file.exists("data")){
  dir.create("data");
}

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists("data/activity_monitoring_data.zip")) {
  download.file(fileUrl, destfile = "./data/activity_monitoring_data.zip")
}

if (!file.exists("./data/activity.csv")) {
  unzip("./data/activity_monitoring_data.zip",exdir="./data")
}

data <- read.csv("./data/activity.csv")

```

## What is mean total number of steps taken per day?

aggregate the steps for each day and the frequency for the total steps per day in a histogram
```{r}
StepsPerDay <- aggregate(steps ~ date, data=data,  FUN=sum, na.rm=TRUE)
png(filename="figure/stepsPerDay.png")
histogram <- hist(StepsPerDay$steps)
dev.off()
histogram <- hist(StepsPerDay$steps)
```

```{r maxAndMean, result=FALSE}
maxSteps <- max(StepsPerDay$steps)
meanSteps <- mean(StepsPerDay$steps)
```

The maximum number of steps in a day is `r maxSteps`
The mean number of steps in a day is `r meanSteps`

## What is the average daily activity pattern?

This figure shows the average steps per 5 minute interval.
```{r}
StepsPerInterval <- aggregate(steps ~ interval, data=data, FUN=mean, na.rm=TRUE)
png(filename="figure/stepsPerInterval.png")
plot <- plot(StepsPerInterval$interval,StepsPerInterval$steps,type="l")
dev.off()
plot <- plot(StepsPerInterval$interval,StepsPerInterval$steps,type="l")

maxsteps <- which.max( StepsPerInterval$steps )
maxsteps

```

The interval with de maximum avarage number of steps is interval `r maxsteps`.

## Imputing missing values

```{r computeMissing,echo=FALSE}
missing <- length(which(is.na(data$steps)))

```

The total number of missing values is `r missing`

create dataset where missing values are replaced bij the avarage of the 5 minute interval
```{r}
library(dplyr)
#StepsPerInterval <- aggregate(steps ~ interval, data=data, FUN=mean, na.rm=TRUE)
#filled <- mutate(data, steps=is.na(steps),StepsPerInterval$steps[interval],steps)
filled=data

```

```{r}
library(dplyr)
StepsPerDayFilled <- aggregate(steps ~ date, data=filled,  FUN=sum, na.rm=TRUE)
png(filename="figure/stepsPerDayFilled.png")
histogram <- hist(StepsPerDayFilled$steps)
dev.off()
histogram <- hist(StepsPerDayFilled$steps)
```

```{r echo=FALSE}
maxFilled <- max(StepsPerDayFilled$steps)
meanFilled <- mean(StepsPerDayFilled$steps)
```

The maximum number of steps per day replacing the not available values with the avarage of the dataset for the interval is `r maxFilled`
The avarege number of steps per day replacing the not available values with the avarage of the dataset for the interval is `r meanFilled`

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(ggplot2)
library(plyr)
Sys.setlocale("LC_TIME", "English")

weekOrWeekend <-  mutate(data,group = ifelse (weekdays(as.Date(data$date),TRUE) %in% c("Sat ","Sun"),"weekend","weekday"))
aggregated <- ddply(weekOrWeekend,c("group","interval"),summarize,steps=mean(steps,na.rm=TRUE))

img <- qplot(interval,steps
             ,data=aggregated
            ,facets=(group~.)
            ,geom=c("line")
            ,main="Avarage number of steps in a 5 minute interval for weekdays and weekend"
            ,xlab="Interval"
            ,ylab="Number of steps"
            )
png(filename="figure/avgStepsInWeekOrWeekend.png")
print(img)
dev.off()
print(img)

```
